\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=0.75in]{geometry}
\usepackage[dvipsnames]{xcolor} % adding colour
\usepackage{amsmath} % adding math equations
\usepackage{natbib} % adding BibTeX support
\usepackage{hyperref}



\begin{document}

	\begin{center}
    
    	% MAKE SURE YOU TAKE OUT THE SQUARE BRACKETS
    
		\LARGE{\textbf{STA380 Project Proposal}} \\
        \vspace{1em}
        \Large{Bootstrap Estimation of Boston Housing Data} \\
        \vspace{1em}

        \normalsize\textbf{Kangyi Huang, Qi Wang, Dongshuo Yang, Jiawei Cao} \\
        \normalsize{xennia.huang@mail.utoronto.ca} \\
        \normalsize{jackieqi.wang@mail.utoronto.ca} \\
        \normalsize{jimmyyang.yang@mail.utoronto.ca} \\
        \normalsize{jiawei.cao@mail.utoronto.ca} \\

	\end{center}

    	\section{Project Topic}
         Bootstrap Estimation of Simple Linear Regression for the Boston Housing Dataset

        \section{Simulation vs. Dataset}
        The project will be using the Boston Housing Dataset, from Harrison and Rubinfeld, 1978, 
        ``Hedonic prices and the demand for clean air” Journal of Environmental Economics and Management. 
        We export the dataset to boston.csv and include it in the submission. We export the dataset to 
        boston.csv and include it in the submission.
        
        \section{Project Details}
        This project be using resampling method, Bootstrap estimation, to estimate simple linear regression
        coefficients (intercept and slope) using the Boston Housing Dataset, which provides 14 potential
        predictors against the housing price in Boston. The result of the Bootstrap sampling distribution
        will be used for contrasting and analyzing with the OLS estimates from the full dataset,
        which serve as reference values for comparison. Bias and standard error of bootstrap estimates are
        assessed relative to these values.
        
        \bigskip %
        \noindent Details of implementation:
        \begin{itemize}
            \item Original data: 506 cases
            \item Bootstrap sample size: 506(same as original)
            \item Number of repetition: 1000
            \item Statistic: User-selected descriptive statistics on chosen variables (with default examples provided below)
        \end{itemize}
        The following outputs will be provided:
        \begin{itemize}
            \item	Bootstrap Sampling Distributions’ histogram summary:
                    \begin{itemize}
            \item	Histogram of beta 1 with OLS reference line
              \begin{itemize}
                \item	A histogram showing the bootstrap estimation distributions of the size of resampling of the slopes compared with the OLS slope. 
              \end{itemize}
              \item	Histogram of bootstrap estimation for beta 0:
            \begin{itemize}
                \item	A histogram showing the bootstrap estimation distributions of the size of resampling of the intercepts, compared with the OLS Intercept.
              \end{itemize}
                    \end{itemize}

            
            \item Numerical Summary of Bootstrap Estimates 
             \begin{itemize}
            \item Which will be provided with the following:
            \begin{itemize}
                \item	bootstrap mean
                \item	bootstrap SE
                \item	Bias (relative to the OLS)
                \item	MSE = variance + Bias$^2$
              \end{itemize}
              \item We also provided the Bias-variance tradeoff to displace the difference between them more intuitively
            \end{itemize}
            \item	Boxplot of Bootstrap vs. OLS Confidence Intervals:
            \begin{itemize}
                \item	Visual comparison of the given alpha confidence interval derived from Bootstrap versus the OLS distribution.
              \end{itemize}
        \end{itemize}

        \section{User Inputs (Shiny Components)}
Users will be able to modify the following:
\begin{enumerate}
  \item Setting the seed
  \item Select the following features to predict:
  \begin{itemize}
    \item Any variable (e.g., MEDV, LSTAT, RM, CRIM, AGE, NOX, etc.)
    \item One or more descriptive statistics (e.g., median, mean, sd, IQR, etc.)
    \item Optional: Pearson correlation between any two selected variables (e.g., corr(MEDV, LSTAT), corr(MEDV, RM))
  \end{itemize}
  \item Select the information presented:
  \begin{itemize}
    \item[-] Bootstrap Sampling Distributions' histogram summary
    \item[-] Numerical Summary of Bootstrap Estimates
    \item[-] Boxplot of Bootstrap vs.\ OLS Confidence Intervals
  \end{itemize}
  \item For the plots, be able to modify the colours.
\end{enumerate}



\section*{AI Statement}
There is no AI utilization for each part in this proposal.

\nocite{*}
\bibliographystyle{abbrvnat}
\bibliography{bibliography}
\vspace{0.8cm}

\end{document}

