---
title: "Bootstrap Estimation for Boston Housing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bootstrap Estimation for Boston Housing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 7,
  fig.height = 6
)
```

This example demonstrates how to perform ***bootstrap estimation for Simple Linear Regression (SLR)*** using the Boston Housing Dataset.

We use the response variable `medv` and the predictor variable `lstat` as an example. We will generate a linear regression model first and then use bootstrap resampling to estimate the sampling variability, bias, and confidence intervals of the intercept and slope.

```{r setup}
data <- read.csv("BostonHousing.csv")
source("main.R")

# Example variables
respond <- "medv"
predictor <- "lstat"
```

The analysis is implemented through the following functions:  
- `ols_estimators(data, predictor, respond)`  
- `bootstrap_slr_summary(data, R, seed, predictor, respond)`  
- `bootstrap_slr(boot_slr, ols_slr)`  
- `bootstrap_slr_ci(boot_slr, ols_slr, level)`  
- `plot_boot_hist(boot_slr, ols_slr, term = c("intercept","slope"))`    
- `plot_ci_box(ci_list, term = c("intercept","slope"))`  
- `plot_lr_bootstrap_scatter(ci_list, term = c("intercept","slope"))`  


## Data

The Boston Housing dataset contains housing and neighborhood variables for census tracts in the Boston area [@harrison1978hedonic].

### Dataset preview

```{r}
dim(data)
names(data)
head(data, 5)
```

### Basic summary for the two variables used

```{r}
summary(data[, c(predictor, respond)])
```

### IQR Visualization: Predictor vs Response

```{r}
plot_iqr_boxplot(data, predictor, respond)
```

### Scatterplot of response vs predictor

```{r}
plot(data[[predictor]], data[[respond]],
     xlab = predictor,
     ylab = respond,
     main = "Boston Housing: medv vs lstat")
```

### Correlation (Pearson) and p-value

```{r}
cor_res <- calculate_correlation(data, predictor, respond)
cor_res
```

## Simple Linear Regression

### OLS estimation

```{r}
ols_slr <- ols_estimators(data, predictor = predictor, respond = respond)
ols_slr
```

We use standard diagnostic plots to check:

1) **Linearity**  
2) **Independence**  
3) **Homoscedasticity**  
4) **Normality of residuals**  

```{r}
lr_model <- lm(reformulate(predictor, respond), data = data)
par(mfrow = c(2, 2))
plot(lr_model)
par(mfrow = c(1, 1))
```

### Residuals vs predictor (extra check)

```{r}
resid_vals <- resid(lr_model)
plot(data[[predictor]], resid_vals,
     xlab = predictor,
     ylab = "Residuals",
     main = "Residuals vs Predictor")
abline(h = 0, lwd = 2, lty = 2)
```

### OLS fitted line on the scatterplot

```{r}
plot(data[[predictor]], data[[respond]],
     xlab = predictor,
     ylab = respond,
     main = "SLR Fit: OLS line on data")
abline(lr_model, lwd = 2, lty = 2)
```

## Bootstrap for Simple Linear Regression

### Bootstrap estimation

We run bootstrap with `R = 1000` resamples.

```{r}
boot_slr <- bootstrap_slr_summary(
  data = data,
  R = 1000,
  seed = 123,
  predictor = predictor,
  respond = respond
)
names(boot_slr)
```

### Bootstrap Correlation Distribution: Predictor vs Response

```{r}
plot_bootstrap_correlation(data, predictor, respond)

```

## Comparative Analysis between Two Models

### Bootstrap summary table (bias / variance / MSE)

```{r}
boot_table <- bootstrap_slr(boot_slr, ols_slr)
knitr::kable(boot_table, digits = 4)
```

### Bootstrap histograms with OLS reference

```{r}
plot_boot_hist(boot_slr, ols_slr, term = "intercept")
```

```{r}
plot_boot_hist(boot_slr, ols_slr, term = "slope")
```

### Confidence interval comparison (Bootstrap percentile vs OLS Wald)

```{r}
ci_list <- bootstrap_slr_ci(boot_slr, ols_slr, level = 0.95)

ci_list$boot_perc
ci_list$ols_wald
```

```{r}
plot_ci_box(ci_list, term = "intercept")
```

```{r}
plot_ci_box(ci_list, term = "slope")
```

### Comparison of Bootstrap estimation and OLS estimation

This figure visualizes the joint sampling distribution of the intercept and slope using the Bootstrap method with 1000 resamples.
We can identify the following message:

Parameter correlation: The tilt direction of the point cloud shows the trade-off between the intercept and slope.

Bias magnitude: The arrow length reflects the degree of deviation between the Bootstrap mean and the OLS estimate.

Model stability: The dispersion of the point cloud corresponds to the variance of the parameter estimates.

Estimation reliability: Whether the orange rhombus falls within a high-density region indicates the representativeness of the original estimate.

```{r}
plot_lr_bootstrap_scatter(ols_slr, boot_slr)
```

## AI Statement

In the Vingette.rmd, the Generative AI (ChatGPT) was used to address the data reading issue.   It also helped understand how to structure the bootstrap summary section, especially how to organize the comparison between bootstrap percentile CI and OLS Wald CI in the text and plots.

In the main.R,  AI provide suggestions on how to visualize the comparison of Bootstrap estimation and OLS estimation. The final decision of using the bivariate graph of the joint sample distribution comes from the knowledge in MAT332 and other STA courses.

When debugging the test_bootstrap_boston, prompted to AI with "Why does the R=100 causing error", the ChatGPT indicates that actually, setting R = 100 itself didn't trigger the red fatal error. The red crosses we saw on lines 124 and 128 during tests were definitely due to the insetting of the variable R. This helped with locating the error in "main.R", which setted R = 1000 inside the function, causing R = 100 in the test giving error.

AI also gave me suggestions on plotting the IQR and Bootstrap distribution plots for the correlation coefficient, and it provided a clear template approach along with some ideas for writing the code. Those helped me a lot to write my own code.

All final choices, variable selections, implementations, interpretations, and visualizations were reviewed and edited by the authors.

## Bibliography

::: {#refs}
:::

