---
title: "Bootstrap Estimation for Boston Housing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bootstrap Estimation for Boston Housing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 7,
  fig.height = 6
)
```

```{r setup, include=FALSE}
data <- read.csv("BostonHousing.csv")
source("main.R")

set.seed(123)

# Choose variables for SLR
respond <- "medv"
predictor <- "lstat"
```

# Overview

This vignette demonstrates bootstrap estimation for a **simple linear regression (SLR)** using the Boston Housing dataset.
We model the relationship between the response variable `medv` and the predictor `lstat`, and then use bootstrap resampling
to estimate sampling variability, bias, and confidence intervals for the intercept and slope.

Workflow:
1) fit OLS SLR and obtain standard (Wald) confidence intervals  
2) bootstrap the SLR coefficients and summarize bias/SE/variance/MSE  
3) compare bootstrap percentile CIs to OLS CIs  
4) check standard linear regression assumptions with diagnostic plots  

# Data

The Boston Housing dataset contains housing and neighborhood variables for census tracts in the Boston area (see [@harrison1978hedonic]).

### Dataset shape and preview

```{r}
dim(data)
names(data)
head(data, 5)
```

### Basic summary for the two variables used

```{r}
summary(data[, c(predictor, respond)])
```

### Visualizing IQR for Selected Variables

Example with predictor = "lstat" and respond = "medv"

```{r iqr-boxplot-example, fig.width=8, fig.height=5, message=FALSE}
# Assuming 'data' is already loaded and main.R is sourced

iqr_result <- plot_iqr_boxplot(
  data      = data,
  predictor = "lstat",
  respond   = "medv",
  col_predictor = "lightblue",
  col_respond   = "lightgreen"
)

# Display numerical IQR values in a table
knitr::kable(
  data.frame(
    Variable = c("Predictor (lstat)", "Response (medv)"),
    IQR      = c(iqr_result$iqr_predictor, iqr_result$iqr_respond)
  ),
  digits = 2,
  caption = "Interquartile Range (IQR) for Selected Variables"
)

# Render the boxplot (knitr will display it automatically)
print(iqr_result$plot)
```

### Scatterplot of response vs predictor

```{r}
plot(data[[predictor]], data[[respond]],
     xlab = predictor,
     ylab = respond,
     main = "Boston Housing: medv vs lstat")
```

### Correlation (Pearson) and p-value

```{r}
cor_res <- calculate_correlation(data, predictor, respond)
cor_res
```

# Functions and Usage

All group functions are sourced from `main.R`. Key functions used here include:

- `ols_estimators(data, predictor, respond)`  
- `bootstrap_slr_summary(data, R, seed, predictor, respond)`  
- `bootstrap_slr(boot_slr, ols_slr)`  
- `bootstrap_slr_ci(boot_slr, ols_slr, level)`  
- `plot_boot_hist(boot_slr, ols_slr, term = c("intercept","slope"))`  
- `plot_ci_box(ci_list, term = c("intercept","slope"))`

## OLS estimation

```{r}
ols_slr <- ols_estimators(data, predictor = predictor, respond = respond)
ols_slr
```

### OLS fitted line on the scatterplot

```{r}
lr_model <- lm(reformulate(predictor, respond), data = data)

plot(data[[predictor]], data[[respond]],
     xlab = predictor,
     ylab = respond,
     main = "SLR Fit: OLS line on data")
abline(lr_model, lwd = 2, lty = 2)
```

# Bootstrap + Linear Regression Assumptions

## Bootstrap estimation

We run bootstrap with `R = 1000` resamples (see [@davison1997bootstrap; @boot2025]).

```{r}
boot_slr <- bootstrap_slr_summary(
  data = data,
  R = 1000,
  seed = 123,
  predictor = predictor,
  respond = respond
)
names(boot_slr)
```

### Bootstrap summary table (bias / variance / MSE)

```{r}
boot_table <- bootstrap_slr(boot_slr, ols_slr)
knitr::kable(boot_table, digits = 4)
```

### Bootstrap histograms with OLS reference

```{r}
plot_boot_hist(boot_slr, ols_slr, term = "intercept")
```

```{r}
plot_boot_hist(boot_slr, ols_slr, term = "slope")
```

### Confidence interval comparison (Bootstrap percentile vs OLS Wald)

```{r}
ci_list <- bootstrap_slr_ci(boot_slr, ols_slr, level = 0.95)

ci_list$boot_perc
ci_list$ols_wald
```

```{r}
plot_ci_box(ci_list, term = "intercept")
```

```{r}
plot_ci_box(ci_list, term = "slope")
```

## Linear regression assumptions (diagnostics)

We use standard diagnostic plots to check:

1) **Linearity**  
2) **Independence**  
3) **Homoscedasticity**  
4) **Normality of residuals**  

```{r}
par(mfrow = c(2, 2))
plot(lr_model)
par(mfrow = c(1, 1))
```

### Residuals vs predictor (extra check)

```{r}
resid_vals <- resid(lr_model)
plot(data[[predictor]], resid_vals,
     xlab = predictor,
     ylab = "Residuals",
     main = "Residuals vs Predictor")
abline(h = 0, lwd = 2, lty = 2)
```

# Bibliography

::: {#refs}
:::

# AI Statement

(Place your final AI statement here according to your course policy.)